---
editor_options: 
  markdown: 
    wrap: 72
---

## Montecarlo Methods in R

### Chapter 1: R

#### Packages in R
```{r}
rm(list =ls())
```


```{r}
#library() function show the available packages
library()
```

```{r}
#Packages are installed with install.packages("package name")
#Packages are called with the library function()
library(combinat)
library(datasets)
library(MASS)
library(coda)
library(mcsm)
```

#### Objects

```{r}
#Function c() concatenates terms together, it can concatenate number
#as well as other vectors
a<-c(10, 20, 20, 10)
#Function str() show the structure of an object
a; str(a)
```

Decimal numbers should be encoded with a dot, character strings in
quotes " ", and logical values with the character strings TRUE and
FALSE. Missing values are encoded with the string NA.

```{r}
#Boolean operators for equality is ==
a == 20
```

\> a[1] display the first element of a\
b=a[2:4] build the numeric vector b of dimension 3 with elements 5.6, 1,
4\
\> d=a[c(1,3,5)] build the numeric vector d of dimension 3 with elements
5, 1, -5\
\> 2\*a multiply each element of a by 2 and display the result\
\> b%%3 provides each element of b modulo 3\
\> d%/%2.4 computes the integer division of each element of d by 2.4\
\> e=3/d build the numeric vector e of dimension 3 and elements 3/5, 3,
-3/5\
\> log(d\*e) multiply the vectors d and e term by term and transform
each term into its natural logarithm\
\> sum(d) calculate the sum of d\
\> length(d) display the length of d\
\> t(d)e transpose d, the result is a row vector\
\> t(d)%\*%e scalar product between the row vector t(b) and the column
vector e with identical length\
\> t(d)\*e elementwise product between two vectors with identical
lengths\
\> g=c(sqrt(2),log(10)) build the numeric vector g of dimension 2 and
elements sqrt(2), log(10)\
\> e[d==5] build the subvector of e that contains the components e[i]
such that d[i]=5\
\> a[-3] create the subvector of a that contains all components of a but
the third.\
\> is.vector(d) display the logical expression TRUE if a vector and
FALSE else

```{r}
#Functions designed for vectors
rm(a)
a <- c(1, 2, 3)
print("sample():")
sample(a)
print("permn():")
permn(a)
print("order():")
order(a)
print("sort():")
sort(a, decreasing = T)
```

```{r}
names(a) = NULL
names(a)
names(a) <- c("1", "Two", "Tres")
names(a)
```

#### Classes: matrix, array, factor

```{r}
#seq function takes (from, to, by, length)
x <- seq(0, 10, by=2)
x
x <- seq(1, 10, length.out=19)
x
```

``` r
> x=matrix(vec,nrow=n,ncol=p)
```

Creates an n x p matrix from a vector of dimension np.

```{r}
a <- c(1:12)
m <- matrix(a, 3, 4, byrow=F)
n <- matrix(a, 4, 3, byrow=T)

m
m[1,]
n
matrix(a, 5, 5)
#Note that the vector can be of different size than np and it will
#either be cut off or have to restart
```


#### Matrix manipulations

```{r}
#Term by termn multiplication
m * m
#Matrix product
m %*% n
#Vector from the diagonal
diag(m)
#It can also create a matrix with a given diagonal
diag(1:5)
#Product with the transpose
crossprod(m, m)
```

\> x1=matrix(1:20,nrow=5) build the numeric matrix x1 of dimension 5 × 4
with first row 1, 6, 11, 16\
\> x2=matrix(1:20,nrow=5,byrow=T) build the numeric matrix x2 of \>
dimension 5 × 4 with first row 1, 2, 3, 4\
\> a=x3%*%x2 matrix summation of x2 and x3\
\> x3=t(x2) transpose the matrix x2\
\> b=x3%*%x2 matrix product between x2 and x3, with a check of the \>
dimension compatibility\
\> c=x1\*x2 term-by-term product between x1 and x2\
\> dim(x1) display the dimensions of x1\
\> b[,2] select the second column of b\
\> b[c(3,4),] select the third and fourth rows of b\
\> b[-2,] delete the second row of b\
\> rbind(x1,x2) vertical merging of x1 and x2\
\> cbind(x1,x2) horizontal merging of x1 and x2\
\> apply(x1,1,sum) calculate the sum of each row of x1\
\> as.matrix(1:10) turn the vector 1:10 into a 10 × 1 matrix\

#### Factor

It is a vector of integers used for discrete classification of the
components of other vectors with the same length.

\> state=c("tas","tas","sa","sa","wa") create a vector with five values\
\> statef=factor(state) distinguish entries by group\
\> levels(statef) give the groups\
\> incomes=c(60,59,40,42,23) create a vector of incomes\
\> tapply(incomes,statef,mean) average the incomes for each group\
\> statef=factor(state, + levels=c("tas","sa","wa","yo")) define a new
level with one more group than observed\
\> table(statef) return statistics for all levels 

#### Classes: list and data.frame

A list is a collection of arbitrary objects known as components.

```{r}
a<-list(name_1=1,name_n="object_n")
a$name_1
a$name_n
```

Components have names associates with them. Lists can also takes other
lists as components as well as any other R objects.

```{r}
li=list(num=1:5,y="color",a=T)
li

a=matrix(c(6,2,0,2,6,0,0,0,36),nrow=3)
res=eigen(a,symmetric=T)
names(res)
res$values
res$vectors

a
res$vec%*%diag(res$val)%*%t(res$vec)
```

#### data.frame

A data frame is a list whose elements are possibly made of differing
modes and attributes but have the same length.

\> v1=sample(1:12,30,rep=T) simulate 30 independent uniform ranfom
variables on {1,2,...,12}\
\> v2=sample(LETTERS[1:10],30,rep=T) simulate 30 independent uniform
random variables on {a, b, ...., j}\
\> v3=runif(30) simulate 30 independent uniform random variables on [0,
1]\
\> v4=rnorm(30) simulate 30 independent realizations from a standard
normal distribution\
\> xx=data.frame(v1,v2,v3,v4) create a data frame\

### Probability distributions  
```{r}
x <- rnorm(20)
x
print(" ")
y <- 3*x + 5 + rnorm(20, sd=0.3)
y

reslm <- lm(y~x)
summary(reslm)
```

### Statistics  
```{r}
#rnorm(n, mean=0, sd=1)
x <- rnorm(20, 0, 0.5)
x
print("")
var(x)
sd(x)
median(x)
mean(x)
quantile(x)
print("")
summary(x)
print("")
t.test(x)
```
```{r}
x <- seq(-3,3,le=5)
x
y <- 2 + 4*x + rnorm(5)
y
a <- lm(y~x)
a
print("")
summary(a)
```

### Graphing
```{r}
par()
```

```{r}
x <- rnorm(100, 0, 1)
y <- rnorm(100, 1, 0.5)
X <- matrix(c(x, y), 100)
hist(X[,1])
hist(X[,2])
```
```{r}
x <- seq(0, 10, len=25)
y <- x^2 + (1/10) * x * rnorm(25, 0, 2)
plot(x, y, main = "x^2 plot with err")
```

### Functions  

```
name=function(arg1[=expr1],arg2[=expr2],...) {
expression
...
expression
value
}
```

#### Example
```{r}
sqrnt = function(y) {
  x=y/2
  while (abs(x*x-y) > 1e-10) 
    x=(x+y/x)/2
  x
}
sqrnt(10)
```
The expressions used in a function rely on a syntax that is quite similar to
those of other programming languages, with conditional statements such as  
if (expres1) expres2 else expres3  
where expres1 is a logical value, and loops such as  
for (name in expres1) expres2  
and  
while (expres4) expres2  


### The mcsm package
```{r}
demo(Chapter.1)
```




