---
editor_options: 
  markdown: 
    wrap: 72
---

## Random variable generation

```{r}
rm(list =ls())
```

```{r}
library(combinat)
library(datasets)
library(MASS)
library(coda)
library(mcsm)
```

#### Uniform simulation

```{r}
# runif(N, min, max)
a <- runif(10000, min = 2, max = 5)
hist(a)
```

```{r}
N = 10^4
b <- runif(N)
b1 = b[-N]
b2 = b[-1]
par(mfrow = c(1,3))
hist(b)
plot(b1, b2)
acf(b)
```

We can change the seed for runif() by running set.seed(int) before the
function.


#### The inverse transform
If X has density f and cdf F, we have:
$$ 
F(x) = \int_{-\infty}^x f(t) dt
$$
If $U = F(X)$ then U is a random variable distributed from a uniform $U(0, 1)$.  
```{r}
N = 10^4
U = runif(N)
X = -log(U)
Y = rexp(N)
par(mfrow=c(1,2))
hist(X, main = "Exp from Uniform")
hist(Y, main = "Exp from R")
```  

#### Transformation methods  
When a distribution with density f is linked in a relatively simple way to
another distribution that is easy to simulate, this relationship can often be exploited to construct an algorithm to simulate variables from f.  
For example, to generate χ 26 random variables, we could
use the R code  
```{r}
U = runif(3 * 10^4)
U = matrix(U, nrow = 3)
X = -log(U)
X = 2 * apply(X, 2, sum)
```

#### Discrete distributions  
Here is an R code that can be used to generate Poisson random
variables for large values of lambda. The sequence t contains the integer values in the range around the mean.  
```{r}
N =  10^4; lambda = 100
spread = 3*sqrt(lambda)
t = round(seq(max(0, lambda - spread), lambda + spread, 1))
prob = ppois(t, lambda)
for (i in 1:N)
{
  u = runif(1)
  X[i] = t[1] + sum(prob < u)
}
```

#### Mixture representation  
It is sometimes the case that a probability distribution can be naturally represented as a mixture distribution; that is, we can write it in the form
$$
f(x) = \int_Y g(x|y)p(y)dy \quad or \quad f(x) = \sum_{i\in Y}p_if_i(x) 
$$  
Example: If X is a negative binomial random variable, X ∼ N eg(n, p),
then X has the mixture representation $X|y\~P(y) \quad and \quad Y\~G(n, \beta)$, where $\beta = (1-p)/p$.  
```{r}
N = 10^4
n = 6; p = 0.3
y = rgamma(N, n, rate = p/(1-p))
x = rpois(N, y)
hist(x, main="", freq=F, breaks = 40, col="grey")
lines(1:50, dnbinom(1:50, n, p), lwd = 2, col="sienna")
```



